Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean throughout the United States, and from Mexico, France, and Australia

### 2017-06-28 doc/manuscript/manuscript.Rmd
Added notes

@attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the US population.
Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic].
Microsatellite analysis revealed little host- specificity or genetic-isolation of *S. sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences.
Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table 1).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table 1).
The bean nurseries were 5 m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the season.
Although the nursery locations were the same over sampling years, sampling fields within a location varied for sampling years.
Sampling was carried out on a 'W' or 'V' shape transect, or randomly.
Sampling year and number of isolates collected were necessarily influenced by incidence and severity of white mold disease.

@@

DAPC is a multivariate method designed to identify and describe clusters of genetically related individuals.
Discriminant analysis maximizes separation between groups while minimizing variation within groups [@jombart2010discriminant].
The `find.clusters()` function in *adegenet* 2.0.1 [@jombart2008adegenet] uses a k-means clustering algorithm to identify the optimum number of clusters.
It was not able to identify a single optimum and identified a range of optimal clusters between 5-15.
Consequently, DAPC was performed several times with different optimal number of clusters to determine whether isolates (N = 318) fell into each cluster have any biological meaning such as belonging to a particular host or came from the same region etc.  the best summaries for the total isolates (N = 318).
Since dbRDA identified region, host, and sampling year as the top three variables influencing microsatellite data, DAPC was also performed on data defined according to each of these variables, separately.
Isolates (N = 318) were assigned to different levels of the three factors (region, host, year) and assessed predefined group size (original number of isolates at each level of the factor), post group size (number of isolates assigned by the DAPC algorithm) and the proportion of the original isolates assigned to the original group after analysis.
This was performed to determine the suitability of explanatory variables (i.e. region, host, and year) in describing the clusters generated.
Clone-corrected isolates sampled in different years from the white mold nurseries (wmn) were tested for population differentiation over time by DAPC analysis.
Only WA, MI, and NE isolates were qualified since we filtered out the states that had at least $\geq$ 7 samples/year with at least two years of data for the analysis.
Lastly, we pooled all isolates, clone corrected them (N = 165) and performed DAPC using the function find.clusters as described above to delineate how many clusters were formed by these unique genotypes.
AMOVA was used to test the significance of clusters resulted by different DAPC analyses using GenAlEx version 6.41 [@peakall2012genalex].
PhiPT which is an analogue of Wright's FST was used to ascertain the significance of populations.
An $F_{st}$ value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
<!-- **(Balloux and Ugon-Moulin, 2002; Wright, 1978)**.   -->

@@

Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity ($H_{exp}$) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
Value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on the , which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
Higher observed  or $I_A$ values would indicate clonally reproducing populations.

@@

A total of 366 isolates were collected from 2003 to 2012 from diseased dry bean plants in 11 states in the US, Australia (2 mainland; 4 Tasmania), Mexico, and France (Fig. 1).
Most isolates came from MI (62), ND (60), WA (59), NE (47), and CO (42).  Other regions in lesser representation were France (24), MX (18), OR (17), CA (16), MN (11), Australia (6), WI (2), ID (1), and NY (1).
Four of the Australian isolates were collected from Tasmania while the other two came from Victoria state.
There were no samples collected in 2006 and 2011.
The highest number of samples were collected in 2005 (N = 63) and lowest in 2012 (N = 4).
Pathogen isolates represented 26 commercial cultivars and breeding lines with some representing unknown dry bean hosts (Fig. 2).
We investigated the population association with host for cultivars where more than 10 isolates were obtained, which included G122, Beryl, Bunsi, Yellow, Pinto, and Merlot.
These six cultivars had a total of 238 *S. sclerotiorum* isolates (of 366 total isolates collected) belonging to 130 MLGs.
Of these MLGs, 53 were singletons, while the rest were obtained from more than one host cultivar; only one private allele on locus 55-4 was identified, belonging to the *S. sclerotiorum* population from the cultivar 'Bunsi'.
The number of unique MLGs recorded from each host is shown in Table 3.
The majority of MLGs reported were found on multiple host cultivars and only a few unique to a cultivar.
For instance, there were only 15 unique MLGs within the population of 68 *S. sclerotiorum* isolates obtained from the 'Beryl' cultivar.

@@

The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively (Fig. 4).
MCGs that were unique to a region are shown in Fig. 5 while MCGs that represented at least 2 or more regions are in Fig 6. 20 MCGs were common in two or more regions and the rest were distributed uniquely in different regions.
ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
The virulence of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
There were six isolates that fell into upper outlier category (virulence > 7.6) and one isolate in the lower outlier range (virulence < 2.3) (box plot results not shown).
The mean virulence of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
ANOVA of these 10 MCGs indicated significant difference for virulence levels at P <0.0001.
Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the virulence levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.

@@

Populations from different regions were compared with AMOVA, where differentiation was low among regions (PhiPT  = 0.11; P = 0.001).
Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P $\leq$ 0.001.
The PhiPT values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
The CA population showed evidence of sexual recombination ( = 0.008; P = 0.27, Ia = 0.08; P = 0.3) while the  value of AU was 0.12 with a high P of 0.05 (Ia = 0.70; P = 0.05), whereas all other populations showed linkage disequilibrium for loci tested indicating high possibility of clonal reproduction.

@@

AMOVA revealed genetic differences among isolates sampled in different years (DAPC biplot not shown).
The PhiPT was a low value of 0.07 with P = 0.001.
Isolates from 2012 were not included in the analysis since there were only 4 samples.
Samples for other years ranged from 19 in 2009 to 63 in 2005.
Sampling year 2003 to 2005 and 2009 were different across every Pairwise comparison.
None of the above sampling variables (i.e. region, year, host) seemed to influence distinct molecular composition of *S. sclerotiorum* populations since pre-defined (i.e. region, host, and year) isolates and isolates that were assigned into post groups by DAPC varied largely (Figs. 8A, 8B and 9).
However, assignment of isolates by DAPC analysis of regions did show some promising results for MX and CA isolates (Fig. 8A).
The original group size of each region and DAPC defined group sizes for different regions are given in Table 6.

@@

We tested the clone corrected data for evidence of sexual outcrossing within *S. sclerotiorum* populations in different regions and different years/region.
CA showed evidence of sexual outcrossing (Ia = 0.08, P.Ia = 0.317,  = 0.008, P. = 0.317) while all other regions implied clonal propagation (data not shown).
Five populations showed evidence of sexual outcrossing: CA in 2004, FR in 2004 and 2005, OR in 2003, and WA in 2008.
However, after clone-correction these populations had 6-7 isolates/population (Table 7) and therefore, may not be sufficient to make inferences about reproductive mode.
Only exception was the WA_2008 population, which consisted of 12 isolates.
Within sampling locations white mold nurseries (wmn) were tested for linkage disequilibrium since same wmn fields were used for sampling over different years.
Only CA wmn showed evidence of sexual reproduction ( = 0.008; P = 0.29) while all other wmn locations suggested clonal reproduction.
The DAPC analyses of WA, and MI isolates from same wmn fields over different years are in Fig. 10.
Since *S. sclerotiorum* isolates are predominantly clonal, we wanted to explore whether there was any introduction of new isolates from other fields or if there were evidence of sexual reproduction.
The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest (PhiPT $\geq$ 0.206, P = 0.001).
The 2003 and 2005 isolates were genetically similar (PhiPT = 0.056, P = 0.129).
MI isolates from 2008 and 2003 (Fig. 10B) showed significant difference (PhiPT = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates (PhiPT $\leq$ 0.078, P $\geq$ 0.076).
NE wmn isolates sampled in 2004 and 2008 were genetically similar (PhiPT = 0.042, P = 0.179) (DAPC results not shown).
The standardized index of association of wmn isolates from each region (i.e.
MI, WA, and NE) for above mentioned sampling years suggested clonal reproduction ( $\geq$ 0.05, P $\leq$ 0.002).

@@

Our study did not find any multilocus genotype/s common across all or majority of sampling geographies.
The most widespread and common genotype (MLG 25, Supplemental Table 1) was found only in five states of the USA with majority of isolates confined to ND.
However, we did find some MLGs of France and Australian isolates to be present in the USA samples.
Some studies on *S. sclerotiorum* did not detect any shared MLGs between geographically distant populations.
@attanayake2013sclerotinia were unable to detect any shared microsatellite MLGs in *S. sclerotiorum* populations from infected canola fields in China and USA.
Isolate collections were limited to a single county in China whereas USA isolates were from ND state only.
This may have limited the chance of discovering genotypes common to both regions.
@atallah2004high examined the population structure of *S. sclerotiorum* in different potato fields of Columbia basin and reported only 15 isolates shared the most common microsatellite haplotype in a sample of 167.
Similarly, several other studies did not detect common fingerprints of *S. sclerotiorum* isolates in geographically distant populations: Australia and North America [@ekins1999genetic];  California, Washington, and Ontario [@malvarez2007new]; Canada, North Carolina, and Louisiana [@cubeta1997clonality].

@@

Interestingly, all Mexican genotypes were unique to that region.
On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material.
Surprisingly, France and Australian genotypes were also isolated in the USA.
MLG 4 (see supplemental Table 1) was recorded from France, Australia and MN while France and USA isolates had four common MLGs.
Similarly, five MLGs were recorded common to both Australia and USA isolates.
However, there were six unique genotypes in clone-corrected France population (N= 6/14) and three genotypes in CA population (N = 3/15).
MLGs from Australian region, our samples were limited to 6, which is not a sufficient number to make inferences.
Our results indicated that different continents were not a defining factor for microsatellite genotypes of *S. sclerotiorum* isolated from dry bean.
This may be due to involuntary movement of sclerotia by human activities and probably because of recent speciation of *S. sclerotiorum*. 
A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor [@carbone1993ribosomal].
Few multilocus genotypes were unique to each sampling state of the USA but most of them were common among several states (Supplemental Table 1).
For instance, Michigan state from which we had the largest collection of isolates (N = 62) consisted of 43 MLGs.
Of these, 32 MLGs were present in at least one of the 5 states of NE, ND, WA, CO, MN; 4 MLGs were present in France isolates; and 11 were unique to MI sites (see Supplemental Table 1 for unique MLGs in different regions).
When selecting *S. sclerotiorum* isolates for screening dry bean germplasm for disease resistance, plant breeders may find above information useful.
A study on genetic diversity of *S. sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance [@aldrich-wolfe2015genetic].
Though our isolates didn't come from different crops, we had isolates from many resistant and susceptible dry bean cultivars, which also didn't show genetic differentiation according to AMOVA.
It seems Mexican pathogen isolates are different from the rest and warrants further investigation using a larger population.
Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region.
Geographic distance between different states of the USA seem not a barrier for gene transfer of *S. sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state.
The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
$H_{exp}$ of MX was 0.26 and indicated less exposure of sampling sites to outside inocula.
Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001,  = 0.37, P. = 0.001).
This may have been another reason for low gene diversity.

DAPC analysis by region followed by AMOVA corroborated the uniqueness of Mexican genotypes by grouping them away from the rest of the isolates (Fig. 8).
Other populations based on geography were unable to group in separate clusters except for the Californian population.
We tested whether genotypes of *S. sclerotiorum* populations were influenced by different dry bean host cultivars and year of sampling by DAPC and AMOVA.
There was no host specificity but year of sampling did show a significant impact on genotypes.
Other studies also reported host identity is no barrier for gene flow of this pathogen [@aldrich-wolfe2015genetic; @clarkson2013population].
@atallah2004high revealed statistically significant impact of the collection date of *S. sclerotiorum* isolates from four potato fields in the Columbia Basin on molecular variance.
But there was no impact of the year or location of collection.
We also did not detect location as a significant predictor of molecular variance but year had a significant impact.
Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection.

>>>>

<!-- SEE: Avoid sentences that start with the authors/reference. -->
@attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the US population.
Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic].
Microsatellite analysis revealed little host- specificity or genetic-isolation of *S. sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences.
Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table 1).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table 1).
The bean nurseries were 5m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the season.
Although the nursery locations were the same over sampling years, sampling fields within a location varied for sampling years.
Sampling was carried out on a 'W' or 'V' shape transect, or randomly.
Sampling year and number of isolates collected were necessarily influenced by incidence and severity of white mold disease.

@@

DAPC is a multivariate method designed to identify and describe clusters of genetically related individuals.
Discriminant analysis maximizes separation between groups while minimizing variation within groups [@jombart2010discriminant].
The `find.clusters()` function in *adegenet* 2.0.1 [@jombart2008adegenet] uses a k-means clustering algorithm to identify the optimum number of clusters.
It was not able to identify a single optimum and identified a range of optimal clusters between 5-15.
Consequently, DAPC was performed several times with different optimal number of clusters to determine whether isolates (N = 318) fell into each cluster have any biological meaning such as belonging to a particular host or came from the same region etc.
Since dbRDA identified region, host, and sampling year as the top three variables influencing microsatellite data, DAPC was also performed on data defined according to each of these variables, separately.
Isolates (N = 318) were assigned to different levels of the three factors (region, host, year) and assessed predefined group size (original number of isolates at each level of the factor), post group size (number of isolates assigned by the DAPC algorithm) and the proportion of the original isolates assigned to the original group after analysis.
This was performed to determine the suitability of explanatory variables (i.e. region, host, and year) in describing the clusters generated.
Clone-corrected isolates sampled in different years from the white mold nurseries (wmn) were tested for population differentiation over time by DAPC analysis.
Only WA, MI, and NE isolates were qualified since we filtered out the states that had at least $\geq$ 7 samples/year with at least two years of data for the analysis.
Lastly, we pooled all isolates, clone corrected them (N = 165) and performed DAPC using the function find.clusters as described above to delineate how many clusters were formed by these unique genotypes.
AMOVA was used to test the significance of clusters resulted by different DAPC analyses using GenAlEx version 6.41 [@peakall2012genalex].
$\Phi_{PT}$ which is an analogue of Wright's $F_{st}$ was used to ascertain the significance of populations.
An $F_{st}$ value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
<!-- **(Balloux and Ugon-Moulin, 2002; Wright, 1978)**.   -->

@@

Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity ($H_{exp}$) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
The value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on $\bar{r}_d$, which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
Higher observed $\bar{r}_d$ or $I_A$ values would indicate clonally reproducing populations.

@@

<!--  ZNK: I think figures 1-4 are a bit redundant with the tables in the text. It might be better to simply remove them. -->
A total of 366 isolates were collected from 2003 to 2012 from diseased dry bean plants in 11 states in the US, Australia (2 mainland; 4 Tasmania), Mexico, and France **(Fig. 1)**.
Most isolates came from MI (62), ND (60), WA (59), NE (47), and CO (42).  Other regions in lesser representation were France (24), MX (18), OR (17), CA (16), MN (11), Australia (6), WI (2), ID (1), and NY (1).
Four of the Australian isolates were collected from Tasmania while the other two came from Victoria state.
There were no samples collected in 2006 and 2011.
The highest number of samples were collected in 2005 (N = 63) and lowest in 2012 (N = 4).
Pathogen isolates represented 26 commercial cultivars and breeding lines with some representing unknown dry bean hosts **(Fig. 2)**.
We investigated the population association with host for cultivars where more than 10 isolates were obtained, which included G122, Beryl, Bunsi, Yellow, Pinto, and Merlot.
These six cultivars had a total of 238 *S. sclerotiorum* isolates (of 366 total isolates collected) belonging to 130 MLGs.
Of these MLGs, 53 were singletons, while the rest were obtained from more than one host cultivar; only one private allele on locus 55-4 was identified, belonging to the *S. sclerotiorum* population from the cultivar 'Bunsi'.
The number of unique MLGs recorded from each host is shown in Table 3.
The majority of MLGs reported were found on multiple host cultivars and only a few unique to a cultivar.
For instance, there were only 15 unique MLGs within the population of 68 *S. sclerotiorum* isolates obtained from the 'Beryl' cultivar.

@@

The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively **(Fig. 4)**. <!-- Maybe add aggressiveness information here. -->
MCGs that were unique to a region are shown in Fig. 5 while MCGs that represented at least 2 or more regions are in Fig 6. 20 MCGs were common in two or more regions and the rest were distributed uniquely in different regions.
ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
The virulence of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
There were six isolates that fell into upper outlier category (virulence > 7.6) and one isolate in the lower outlier range (virulence < 2.3) (box plot results not shown).
The mean virulence of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
ANOVA of these 10 MCGs indicated significant difference for virulence levels at P <0.0001.
Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the virulence levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.

@@

Populations from different regions were compared with AMOVA, where differentiation was low among regions ($\Phi_{PT}$  = 0.11; P = 0.001).
Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P $\leq$ 0.001.
The $\Phi_{PT}$ values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
The CA population showed evidence of sexual <!-- ZNK: There is the sample size caveat here that should be discussed. -->
recombination ($\bar{r}_d$ = 0.008; P = 0.27, $I_A$ = 0.08; P = 0.3) while the $\bar{r}_d$ value of AU was 0.12 with a high P of 0.05 ($I_A$ = 0.70; P = 0.05), whereas all other populations showed linkage disequilibrium for loci tested indicating high possibility of clonal reproduction.

@@

AMOVA revealed genetic differences among isolates sampled in different years (DAPC biplot not shown).
The $\Phi_{PT}$ was a low value of 0.07 with P = 0.001.
Isolates from 2012 were not included in the analysis since there were only 4 samples.
Samples for other years ranged from 19 in 2009 to 63 in 2005.
Sampling year 2003 to 2005 and 2009 were different across every Pairwise comparison.
None of the above sampling variables (i.e. region, year, host) seemed to influence distinct molecular composition of *S. sclerotiorum* populations since pre-defined (i.e. region, host, and year) isolates and isolates that were assigned into post groups by DAPC varied largely (Figs. 8A, 8B and 9).
However, assignment of isolates by DAPC analysis of regions did show some promising results for MX and CA isolates (Fig. 8A).
The original group size of each region and DAPC defined group sizes for different regions are given in Table 6.

@@

We tested the clone corrected data for evidence of sexual outcrossing within *S. sclerotiorum* populations in different regions and different years/region.
CA showed evidence of sexual outcrossing (Ia = 0.08, P.Ia = 0.317, $\bar{r}_d$ = 0.008, P.$\bar{r}_d$ = 0.317) while all other regions implied clonal propagation (data not shown).
Five populations showed evidence of sexual outcrossing: CA in 2004, FR in 2004 and 2005, OR in 2003, and WA in 2008.
However, after clone-correction these populations had 6-7 isolates/population (Table 7) and therefore, may not be sufficient to make inferences about reproductive mode.
Only exception was the WA_2008 population, which consisted of 12 isolates.
Within sampling locations white mold nurseries (wmn) were tested for linkage disequilibrium since same wmn fields were used for sampling over different years.
Only CA wmn showed evidence of sexual reproduction ( = 0.008; P = 0.29) while all other wmn locations suggested clonal reproduction.
The DAPC analyses of WA, and MI isolates from same wmn fields over different years are in Fig. 10.
Since *S. sclerotiorum* isolates are predominantly clonal, we wanted to explore whether there was any introduction of new isolates from other fields or if there were evidence of sexual reproduction.
The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest ($\Phi_{PT}$ $\geq$ 0.206, P = 0.001).
The 2003 and 2005 isolates were genetically similar ($\Phi_{PT}$ = 0.056, P = 0.129).
MI isolates from 2008 and 2003 (Fig. 10B) showed significant difference ($\Phi_{PT}$ = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates ($\Phi_{PT}$ $\leq$ 0.078, P $\geq$ 0.076).
NE wmn isolates sampled in 2004 and 2008 were genetically similar ($\Phi_{PT}$ = 0.042, P = 0.179) (DAPC results not shown).
The standardized index of association of wmn isolates from each region (i.e.
MI, WA, and NE) for above mentioned sampling years suggested clonal reproduction ( $\geq$ 0.05, P $\leq$ 0.002).

@@

Our study did not find any multilocus genotype/s common across all or majority of sampling geographies. <!-- ZNK: I think this statement is a bit misleading. Yes, we didn't find any individual MLGs, but 46% of these are present across at least two populations.  -->
The most widespread and common genotype (MLG 25, Supplemental Table 1) was found only in five states of the USA with majority of isolates confined to ND.
However, we did find some MLGs of France and Australian isolates to be present in the USA samples.
Some studies on *S. sclerotiorum* did not detect any shared MLGs between geographically distant populations.
@attanayake2013sclerotinia were unable to detect any shared microsatellite MLGs in *S. sclerotiorum* populations from infected canola fields in China and USA.
Isolate collections were limited to a single county in China whereas USA isolates were from ND state only.
This may have limited the chance of discovering genotypes common to both regions.
@atallah2004high examined the population structure of *S. sclerotiorum* in different potato fields of Columbia basin and reported only 15 isolates shared the most common microsatellite haplotype in a sample of 167.
Similarly, several other studies did not detect common fingerprints of *S. sclerotiorum* isolates in geographically distant populations: Australia and North America [@ekins1999genetic];  California, Washington, and Ontario [@malvarez2007new]; Canada, North Carolina, and Louisiana [@cubeta1997clonality].

@@

Interestingly, all Mexican genotypes were unique to that region.
On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material. <!-- ZNK: Do we have some sort of evidence for this, or is this pure speculation? -->
<!-- 
I'm not sure what the point of this paragraph is.
I believe I can put the discussion of the graph results here. 
-->
Surprisingly, France and Australian genotypes were also isolated in the USA.
MLG 4 (see supplemental Table 1) was recorded from France, Australia and MN while France and USA isolates had four common MLGs.
Similarly, five MLGs were recorded common to both Australia and USA isolates.
However, there were six unique genotypes in clone-corrected France population (N= 6/14) and three genotypes in CA population (N = 3/15).
MLGs from Australian region, our samples were limited to 6, which is not a sufficient number to make inferences.
Our results indicated that different continents were not a defining factor for microsatellite genotypes of *S. sclerotiorum* isolated from dry bean.
This may be due to involuntary movement of sclerotia by human activities and probably because of recent speciation of *S. sclerotiorum*. 
A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor [@carbone1993ribosomal].
Few multilocus genotypes were unique to each sampling state of the USA but most of them were common among several states (Supplemental Table 1).
For instance, Michigan state from which we had the largest collection of isolates (N = 62) consisted of 43 MLGs.
Of these, 32 MLGs were present in at least one of the 5 states of NE, ND, WA, CO, MN; 4 MLGs were present in France isolates; and 11 were unique to MI sites (see Supplemental Table 1 for unique MLGs in different regions).
When selecting *S. sclerotiorum* isolates for screening dry bean germplasm for disease resistance, plant breeders may find above information useful.
A study on genetic diversity of *S. sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance [@aldrich-wolfe2015genetic].
Though our isolates didn't come from different crops, we had isolates from many resistant and susceptible dry bean cultivars, which also didn't show genetic differentiation according to AMOVA.
It seems Mexican pathogen isolates are different from the rest and warrants further investigation using a larger population.
Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region. <!-- ZNK: I'm not entirely sold on this conclusion. -->
Geographic distance between different states of the USA seem not a barrier for gene transfer of *S. sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state. <!-- ZNK: I don't know that this adds anything. -->
The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
$H_{exp}$ of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-- ZNK: Does this mean that there's heavy inbreeding here? -->
Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001, $\bar{r}_d$ = 0.37, P.$\bar{r}_d$  = 0.001).
This may have been another reason for low gene diversity.

DAPC analysis by region followed by AMOVA corroborated the uniqueness of Mexican genotypes by grouping them away from the rest of the isolates (Fig. 8).
Other populations based on geography were unable to group in separate clusters except for the Californian population.
We tested whether genotypes of *S. sclerotiorum* populations were influenced by different dry bean host cultivars and year of sampling by DAPC and AMOVA.
There was no host specificity but year of sampling did show a significant impact on genotypes.
Other studies also reported host identity is no barrier for gene flow of this pathogen [@aldrich-wolfe2015genetic; @clarkson2013population].
@atallah2004high revealed statistically significant impact of the collection date of *S. sclerotiorum* isolates from four potato fields in the Columbia Basin on molecular variance.
But there was no impact of the year or location of collection.
We also did not detect location as a significant predictor of molecular variance but year had a significant impact.
Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection. <!-- BSA: Is this true, coauthors? -->
